Rotational Spectral Unmixing of Exoplanets:\\Degeneracies between Surface Colors and Geography

### 2017-03-10 paper/main.tex
latest changes...

\def\memoYF#1{\color{red}[NOTE: {\bf #1}]\color{black}}

@@

\author{
%
Yuka Fujii\altaffilmark{1,2} 
%
Jacob Lustig-Yaeger\altaffilmark{3} 
%
Nicolas Cowan\altaffilmark{4} 
%
}

@@

\affil{$^3$University of Washington, 
  }

\affil{$^4$McGill University, 
  }

@@

Direct imaging is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of the Earth and other planets, as they are observed from an astronomical distance. 
Atmospheric molecules are identifiable through spectral absorption features \citep[e.g.,][]{DesMarais2002} but surface reflectance spectra also affect the spectra, which could be measured through low-resolution spectra, or multi-band photometry \citep[e.g.,][]{Ford2001}.

However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

A key here is to leverage the time variation of the spectrum: the regions that contributes to the scattered light change due to the planetary rotation and  orbital motion, so the time variability can in principle be translated to the heterogeneity of the surface environment.  
\citet{Cowan2009, Cowan2011} performed Principal Component Analysis (PCA) on the observed multi-band photometry of Earth. They found that the number of surface types can be inferred from the number of dominant principle components: (\# of surface types) $\ge $ (\# of principal components) $+ 1$. %) and that the variation pattern is indicative of the surface properties. 
\citet{Fujii2010, Fujii2011} decomposed multi-band photometric data of the Earth as a linear inverse problem, assuming the template reflectance spectra of the known major surface types, and they found that the relative abundance and the longitudinal variation of these surface types are approximately recovered. 
Moreover, by coupling the time variation due to rotation and orbital revolution, a 2-dimensional information of the surface may be retrieved \citep{Kawahara2010, Kawahara2011, Fujii2012}.

@@

The assumed albedo spectra of these surface types in 4 photometric bands are shown in the upper right panel of Figure \ref{fig:mockdata}. 
These 4 photometric bands actually correspond to the 0.4-0.5 $\mu $m, 0.5-0.6 $\mu $m, 0.6-0.7 $\mu $m, and 0.7-0.8 $\mu $m, respectively, but they will be simply called by the band indices unless otherwise noted . 
The albedo spectrum for ocean is based on \citet{Mclinden1997}, 
and the data for soil and vegetation are taken from ASTER spectral library\footnote{\url{https://speclib.jpl.nasa.gov/}. 
Specifically, we adopt  ``Brown to dark brown sand'' for ``soil'', and ``Grass'' for the ``vegetation''}. 
The scattering phase function by the surface is assumed to obey the Lambert law, i.e., the radiance is (incident flux)$\times $(albedo)/$\pi$, independent of the direction of the scattering.

@@

The time variability of the apparent covering fraction $\fast $ due to the planet's rotation is related to the surface inhomogeneity along the equator. Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int_{{\rm IV}_{i}} f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int_{{\rm IV}_{il}}  f_k (\vec \Omega ) W (\vec \Omega  ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \\
&\approx & \sum_l f_{lk} \frac{ \int_{{\rm IV}_{il}} W (\vec \Omega ) \cdot d\vec \Omega }{ \int_{{\rm IV}}  W (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\equiv & \sum_l  W_{il} f_{lk} \label{eq:Wf}, 
\end{eqnarray}
%%%
where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
Strictly speaking, the approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little across the $l$-th slice for all $k$ and $l$. 
\memoYF{Comment from Nick: is this approximation exact in the zero obliquity, edge-on limit?}
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = \sum _{l,k} W_{il} \, f_{lk} \, s_{kj} \label{eq:d_f_s}
\end{equation}
%%%
where $f_{lk}$ is the average area fraction of $k$-th surface type at $l$-th longitudinal slice, Again, the area fractions at longitudinal slices, $f_{lk}$, cannot be negative and should sum up to unity. Thus, a set of condition similar to Equations (\ref{eq:tilde_s_range})-(\ref{eq:tilde_f_sum}) are imposed:
%%%
\begin{subnumcases}
{}
0 \leq f_{lk} \;\;\; & \mbox{for any $l$, $k$} \label{eq:f_range} \\
\sum_k f_{lk} = 1 & \mbox{for any $l$} \label{eq:f_sum} \\
0 \leq s_{kj} \leq 1 \;\;\; & \mbox{for any $k$, $j$} \label{eq:s_range}
\end{subnumcases}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of Equations (\ref{eq:s_range})-(\ref{eq:f_sum})---this is where \citet{Cowan2013} stood.

@@

This (hyper-)plane can be identified through Principal Component Analysis (PCA) \citep{Cowan2009,Cowan2011}, as PCA extracts the major, mutually-orthogonal axes along which the scatter among the data points are significant. 
Consequently, the number of major principal components (PCs) are $K-1$ \citep{Cowan2011}. 
%Note that the number of principal components In other words, the number of spectrally distinct major surface types ($=K$) is the number of dominant principal components (PCs) plus 1 in theory \citep{Cowan2011}. 
% used Principal Component Analysis (PCA) to 
% The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-the dimensional space is inconvenient for visualization. 
%We thus start by performing principal component analysis (PCA). 
From the mock light curves prepared in Section \ref{s:mockdata}, we extract two dominant PCs through PCA with other components having virtually zero contributions, consistent with the 3 input surface types.  
The PCs extracted from the light curves at $\alpha = 90^{\circ }$ (olive lines in Figure \ref{fig:mockdata}) are presented in the upper panel of Figure \ref{fig:trajectory}. 
In the following, we adopt these two PCs as the axes of the PC plane. 
\memoYF{Comment from Nick: This is true by definition?}
For the later use, we denote these PCs by $V_{nj}$ where $n=1$ or $2$  corresponding to the 1st and 2nd PC. 
Although the PCs from other light curves with different phase angles are not necessarily same as the one shown in the upper panel of Figure \ref{fig:trajectory}, the PC plane itself should be the same plane (under the noiseless condition), because we assume the same 3 surface types. 
Thus in this section we keep using is PCs to discuss other light curves as well.

@@

Additional constraint comes from the condition (\ref{eq:tilde_s_range}). 
In Figure \ref{fig:trajectory} any points in the shadowed region are rejected based on this condition: specifically, the left, right, and bottom boundaries are set by $s_{k,4}> 0$, $s_{k,1}> 0$, and $s_{k,3}> 0$. 
While in this particular example the permitted region happen to have a triangle shape, the shapes can have a different shape depending on the relative configuration of the boundariese\memoYF{In more detail?}. 
Although these limits the region where $\{ {\bf s}_k \}$ can exist,   
this is in general not sufficient to result in a unique solution.

In Figure \ref{fig:trajectory} we show two random triangles that enclose the trajectory; these triangles satisfy the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), so do many others. 
One can have spectrally interesting surface spectra and boring geography (small longitudinal variation in area fractions) {\it or} one can have boring surface spectra (closer to flat) and interesting geography. \memoYF{Need to reconsider this phrasing. }
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
%\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_lon_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types2_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage} 
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l30deg.pdf}
    \end{center}
     \end{minipage}      
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l40deg.pdf}
    \end{center}
     \end{minipage}
    \caption{(From top-left) A: Probability of surface albedo spectra estimated from the light curves of $\alpha = 90^{\circ }$ in the procedure described in Section \ref{ss:guess}. B: same as A but from the light curves of  $\alpha = 135^{\circ }$. C: same as C but from the colors of longitudinal slices.  D: same as A but from the light curves where the geographical map of soil and vegetation are switched. E: same as A but with regularization with $\lambda = 30^{\circ }$. same as A but with regularization with $\lambda = 40^{\circ }$. }
\label{fig:noreg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given that any combinations that enclose the trajectory/-ies can be a solution, the choice of the solution critically depends on the prior probability distribution. 
% Thus, we need to be careful about the implicit assumptions that are made  
With no assumptions or information on the spectral albedo of the surface types or geography, it may be reasonable to assume that any points in the PC plane except for the inhibited regions are equally likely to correspond to a surface type. 
While such an assumption does not seem to work at all to choose the surface types, it does put some constrains. 
This is because of the relative configuration of the trajectory and the permitted region. 
For example, in Figure \ref{fig:trajectory}, in order to enclose the trajectory with three surface types within the permitted region, we must have at least one point near the bottom left corner of the permitted region (white triangle); this is consistent with the fact that one of the input albedo spectra (ocean) resides there.

In order to see this more quantitatively, we make a grid on the PC plane  with an interval of 0.2 and consider all combinations of three grid points. 
Then, we assumed that all of the combinations that enclose the trajectory are equally likely to be a solution, and found the marginalized probability of having a surface type at each location of the PC plane. 
Specifically, we consider the following Bayesean expression and attempt to find the posterior distribution:
%%%
\begin{equation}
P_{\rm posterior} ( t_{kn} | U_{in} ) = P ( U_{in} | t_{kn} ) \cdot P_{\rm prior} (  t_{kn} ) 
\end{equation}
%%%
where $t_{kn}$ represents the coordinates of the $k$-th surface type on the PC plane ($n$ is either 1 or 2, corresponding to PC 1 and PC 2), i.e., 
%%%
\begin{equation}
s_{kj} = \sum_n t_{kn} V_{nj} + \bar d_j . 
\end{equation}
Assuming that any facet has the same prior probability per area of being a surface type, the prior for each $k$ is:
%%%
\begin{eqnarray}
&& P_{\rm prior} (t_{k1}, t_{k2} ) \, dt_{k1} dt_{k2} \\
&& = \left\{
\begin{array}{ll}
\displaystyle \frac{dt_{k1}  dt_{k2}}{\iint _{\rm allowed} d^2{\bf t}_k} & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_s_range})} \\
0 & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
On the other hand, the likelihood is 
%%%
\begin{eqnarray}
&& P ( U_{in} | t_{kn} ) \\
&& \propto \left\{
\begin{array}{ll}
1 \;\;\; & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_range})} \\
0 \;\;\; & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
Although the graphical meaning of conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) is simply to enclose the data, the judgement is numerically made as follows. 
We first find $\fast $ that satisfies condition (\ref{eq:tilde_f_sum}) from
%%%
\begin{equation}
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
= 
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}
\end{equation}
%%%
or,
\begin{equation}
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
=
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}^{-1}
\label{eq:f=ds-1}
\end{equation}
%%%
Note that the last matrix is a $K\times K$ matrix, and thus inverse is found if $\{ {\bf t}_k \}$ forms a triangle (rather than a straight line or a single dot, in which case $\{ {\bf t}_k \}$ is not a solution anyways).
Then. we ask if the resultant $\fast_{ik}$ are within the range of (\ref{eq:tilde_f_range}).

@@

@@

We must take caution here that the actual covering fraction of surface types do not necessarily obey gaussian process. 
In this particular case, it appears that the dominant surface types are better approximated by gaussian processes.... \memoYF{??}

\memoYF{Have to do with regularization on the maps}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EPOXI data}
\label{s:EPOXI}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=\hsize]{raddata_2_PCplane_noreg.pdf}
    \end{center}
    \caption{}
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we apply our analysis procedure to the observed multi-band light curves of the Earth observed by EPOXI mission. 
Specifically we use the same datasets as \citet{Cowan2013}, i.e., the diurnal light curves observed in June of 2008. 
Using two dominant PCs we extracted, we performed the same analysis described in the previous section

Figure \ref{fig:EPOXI} shows the trajectory of the light curve projected onto the PC plane, with the grayed region being the forbidden region. 
While the permitted region appears to be close to a square, it is in fact a heptagon, bounded by 7 inequalities. Overall, the upper boundaries come from the requirement of $s_{kj}<1$ while the lower boundaries comes from the requirement of $s_{kj}<0$.

>>>>

\def\memoYF#1{\color{red}[YF: {\bf #1}]\color{black}}
\def\memoJLY#1{\color{green}[JLY: {\bf #1}]\color{black}}
\def\memoNBC#1{\color{blue}[NBC: {\bf #1}]\color{black}}

@@

\author{
%
Yuka Fujii\altaffilmark{1,2} 
%
Jacob Lustig-Yaeger\altaffilmark{3,4,5} 
%
Nicolas B. Cowan\altaffilmark{6,7} 
%
}

@@

\affil{$^3$Astronomy Department, University of Washington, Box 951580, Seattle, WA 98195, USA}

\affil{$^4$Astrobiology Program, University of Washington, 3910 15th Ave. NE, Box 351580, Seattle, WA 98195, USA}

\affil{$^5$NASA Astrobiology Institute -- Virtual Planetary Laboratory Lead Team, USA}

\affil{$^6$Department of Earth and Planetary Sciences, McGill University, Montreal, Quebec
Canada H3A 0E8}

\affil{$^7$Department of Physics, McGill University, Montreal, Quebec
Canada H3A 0E8}

@@

Direct imaging is expected to play a vital role in characterizing Earth analogs in habitable zones and beyond. 
Substantial work has gone into prediciting detectable features in disk-integrated spectra of the Earth and other planets, as they are observed from an astronomical distance. 
Atmospheric molecules are identifiable through spectral absorption features \citep[e.g.,][]{DesMarais2002} but surface reflectance spectra also affect the spectra, which could be measured through low-resolution spectra, or multi-band photometry \citep[e.g.,][]{Ford2001}. 
However, interpreting disk-integrated colors is not trivial. 
This is particularly true for Earth-like planets that harbor diverse atmospheric and surface characteristics including liquid water, partial cloud cover, continents, and possibly vegetation.

A key here is to leverage the time variation of the spectrum: the regions that contributes to the scattered light change due to the planetary rotation and  orbital motion, so the time variability can in principle be translated to the heterogeneity of the surface environment.  
\citet{Cowan2009, Cowan2011} performed Principal Component Analysis (PCA) on the observed multi-band photometry of Earth. They found that the number of surface types can be inferred from the number of dominant principle components: (\# of surface types) $\ge $ (\# of principal components) $+ 1$. %) and that the variation pattern is indicative of the surface properties. 
\citet{Fujii2010, Fujii2011} decomposed multi-band photometric data of the Earth assuming the template reflectance spectra of the known major surface types, and they found that the relative abundance and the longitudinal variation of these surface types are approximately recovered. 
Moreover, by coupling the time variation due to rotation with phase variation due to orbital revolution, a 2-dimensional information of the surface may be retrieved \citep{Kawahara2010, Kawahara2011, Fujii2012}.

@@

The assumed albedo spectra of these surface types in 4 photometric bands are shown in the upper right panel of Figure \ref{fig:mockdata}. 
These 4 photometric bands actually correspond to the 0.4-0.5 $\mu $m, 0.5-0.6 $\mu $m, 0.6-0.7 $\mu $m, and 0.7-0.8 $\mu $m, respectively, but they will be simply called by the band indices unless otherwise noted. \memoJLY{I think it would be helpful to make the "band \#"-axis wavelength in microns, and then perhaps refer to the band \# in the text or in an upper x-axis. }
The albedo spectrum for ocean is based on \citet{Mclinden1997}, 
and the data for soil and vegetation are taken from ASTER spectral library\footnote{\url{https://speclib.jpl.nasa.gov/}. 
Specifically, we adopt  ``Brown to dark brown sand'' for ``soil'', and ``Grass'' for the ``vegetation''}. 
The scattering phase function by the surface is assumed to obey the Lambert law, i.e., the radiance is (incident flux)$\times $(albedo)/$\pi$, independent of the direction of the scattering. 
In reality, ocean is not a Lambertian scatterer,

\memoJLY{A phase of 135 deg is right in the range that Ty (Robinson et al. 2010; 2011) shows will deviate most from a Lambertian scatterer due the ocean glint...}

@@

The time variability of the apparent covering fraction $\fast $ due to the planet's rotation is related to the surface inhomogeneity along the equator. Approximately, $\fast _{ik}$ may be written as the weighted summation of the area fraction of $k$-th surface type in each of (the finite number $L$ of) longitudinal slices, i.e.,
%%%
\begin{eqnarray}
\fast _{ik} &=& \frac{ \int f_k (\vec \Omega ) \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }{ \int_{{\rm IV}_i}  \cos \theta_0 ({\vec \Omega}) \cdot \cos \theta_1 ({\vec \Omega}) \cdot d\vec \Omega }  \\
&=& \sum_l \frac{ \int f_k (\vec \Omega ) W_i (\vec \Omega  ) \cdot d\vec \Omega }{ \int  W_i (\vec \Omega ) \cdot d\vec \Omega } \\
%&\approx & \sum_l f_{lk} \frac{ \int W_i (\vec \Omega ) \cdot d\vec \Omega }{ \int  W_i (\vec \Omega ) \cdot d\vec \Omega } \label{eq:discretize}\\
&\approx & \sum_l  W_{il} f_{lk} \label{eq:Wf}, \\
W_{il} &\equiv & \frac{ \int  W_i (\vec \Omega  ) \cdot d\vec \Omega_l }{ \int W_i (\vec \Omega )  \cdot d\vec \Omega }
\end{eqnarray}
%%%

where $l$ is the index for longitudinal slices, $f_{lk}$ is the area fraction of the $k$-th surface type in the $l$-th longitudinal slice. 
Strictly speaking, the approximation is valid only when $f_k(\vec \Omega)$ does not change or changes little across the $l$-th slice for all $k$ and $l$. 
\memoNBC{Is this approximation exact in the zero obliquity, edge-on limit?} \memoYF{I don't think so. It is due to pixelization, and does not depend on geometry, I think. ( Perhaps am I not understanding your question properly?) If we suppose the equation were exact, 
%%%
\begin{eqnarray}
f_{lk} &=& \frac{ \int f_k (\vec \Omega ) W_i (\vec \Omega  ) \cdot d\vec \Omega_l }{ \int W_i (\vec \Omega )  \cdot d\vec \Omega_l }\\
\end{eqnarray}
%%%
} 
In the last equation, $W_{il}$ is the weight function for $i$-th epoch and $l$-th longitudinal slice which depends only on the observational geometry. 
As a result,
%%%
\begin{equation}
d_{ij} = \sum _{l,k} W_{il} \, f_{lk} \, s_{kj} \label{eq:d_f_s}
\end{equation}
%%%
where $f_{lk}$ is the average area fraction of $k$-th surface type at $l$-th longitudinal slice, Again, the area fractions at longitudinal slices, $f_{lk}$, cannot be negative and should sum up to unity. Thus, a set of condition similar to Equations (\ref{eq:tilde_s_range})-(\ref{eq:tilde_f_sum}) are imposed:
%%%
\begin{subnumcases}
{}
0 \leq f_{lk} \;\;\; & \mbox{for any $l$, $k$} \label{eq:f_range} \\
\sum_k f_{lk} = 1 & \mbox{for any $l$} \label{eq:f_sum} \\
0 \leq s_{kj} \leq 1 \;\;\; & \mbox{for any $k$, $j$} \label{eq:s_range}
\end{subnumcases}
%%%
Now, the relevant problem is, given $d$, estimate $\{f, s\}$ subject to the constraints of Equations (\ref{eq:s_range})-(\ref{eq:f_sum})---this is where \citet{Cowan2013} stood.

@@

This (hyper-)plane can be identified through Principal Component Analysis (PCA) \citep{Cowan2009,Cowan2011}, as PCA extracts the major, mutually-orthogonal axes along which the scatter among the data points are significant. 
Consequently, the number of major principal components (PCs) are $K-1$ \citep{Cowan2011}. 
%Note that the number of principal components In other words, the number of spectrally distinct major surface types ($=K$) is the number of dominant principal components (PCs) plus 1 in theory \citep{Cowan2011}. 
% used Principal Component Analysis (PCA) to 
% The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-the dimensional space is inconvenient for visualization. 
%We thus start by performing principal component analysis (PCA). 
From the mock light curves prepared in Section \ref{s:mockdata}, we extract two dominant PCs through PCA with other components having virtually zero contributions, consistent with the 3 input surface types.  
The PCs extracted from the light curves at $\alpha = 90^{\circ }$ (olive lines in Figure \ref{fig:mockdata}) are presented in the upper panel of Figure \ref{fig:trajectory}. 
In the following, we adopt these two PCs as the axes of the PC plane. 
\memoNBC{This is true by definition?}
\memoNBC{You mean if you have the same 3 surfaces?}
\memoYF{What I meant was that because we assume the same 3 surface types, the data (apparent albedo) should lie in the same plane determined by these surface types. So although it is not guaranteed that we extract the same PCs from the light curves with different phase angles (because it is determined by the variation pattern of the data), the PC plane itself should be the same. So, if our purpose of doing PCs are to find the PC plane and display it, we could use a consistent PCs to show the trajectories on PC plane. }
For the later use, we denote these PCs by $V_{nj}$ where $n=1$ or $2$  corresponding to the 1st and 2nd PC. 
Although the PCs from other light curves with different phase angles are not necessarily same as the one shown in the upper panel of Figure \ref{fig:trajectory}, the PC plane itself should be the same plane (under the noiseless condition), because we assume the same 3 surface types. 
\memoJLY{I'm skeptical of this claim...Does it break down when the Lambertian assumption is relaxed? This would change the apparent albedo of some surfaces as a function of phase...}
Thus in this section we keep using is PCs to discuss other light curves as well.

@@

Additional constraints come from the condition (\ref{eq:tilde_s_range}). 
In Figure \ref{fig:trajectory} any points in the shadowed region are rejected based on this condition: specifically, the left, right, and bottom boundaries are set by $s_{k,4}> 0$, $s_{k,1}> 0$, and $s_{k,3}> 0$. 
While in this particular example the permitted region happens to be a triangle, shape can differ depending on the relative configuration of the boundaries\memoYF{In more detail?}. 
Although this limits the region where $\{ {\bf s}_k \}$ can exist,   
this is in general not sufficient to result in a unique solution.

In Figure \ref{fig:trajectory} we show two random triangles that enclose the trajectory; these triangles satisfy the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), as do many others. 
One can have spectrally interesting surface spectra and boring geography (small longitudinal variation in area fractions) {\it or} one can have boring surface spectra (closer to flat) and interesting geography. \memoYF{Need to reconsider this phrasing. }
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
%\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_lon_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types2_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage} 
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l30deg.pdf}
    \end{center}
     \end{minipage}      
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l40deg.pdf}
    \end{center}
     \end{minipage}
    \caption{(From top-left) A: Probability of surface albedo spectra estimated from the light curves of $\alpha = 90^{\circ }$ in the procedure described in Section \ref{ss:guess}. B: same as A but from the light curves of  $\alpha = 135^{\circ }$. C: same as A but from the colors of longitudinal slices.  D: same as A but from the light curves where the geographical maps of soil and vegetation are swaped. E: same as A but with regularization with $\Delta_ t = 2$ hr. F: same as A but with regularization with $\Delta_ t = 2.7$ hr. \memoYF{This is a tentative set of panels.} }
\label{fig:noreg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given that any vertices in the PC plane that enclose the trajectory/-ies can be a solution, the choice of the solution critically depends on the prior probability distribution. 
% Thus, we need to be careful about the implicit assumptions that are made  
With no assumptions or information on the spectral albedo of the surface types or geography, it may be reasonable to assume that any points in the PC plane except for the inhibited regions are equally likely to correspond to a surface type. 
%While such an assumption does not seem to work at all to choose the surface types, it does put some constrains. 
While such an assumption does not uniquely specify the colors of all $K$ surface types, it does offer constraints due to the relative light curve trajectory within the permitted region. 
%This is because of the relative configuration of the trajectory and the permitted region. 
For example, in Figure \ref{fig:trajectory}, in order to enclose the trajectory with three surface types within the permitted region, we must have at least one point near the bottom left corner of the permitted region (white triangle); this is consistent with the fact that one of the input albedo spectra (ocean) resides there. \memoJLY{and is a dominant surface type...?}

In order to see this more quantitatively, we make a grid on the PC plane  with an interval of 0.2 and consider all combinations of three grid points. 
Then, we assumed that all of the combinations that enclose the trajectory are equally likely to be a solution, and found the marginalized probability of having a surface type at each location of the PC plane. 
Specifically, we consider the following Bayesean expression and attempt to find the posterior distribution:
%%%
\begin{equation}
P_{\rm posterior} ( t_{kn} | U_{in} ) = P ( U_{in} | t_{kn} ) \cdot P_{\rm prior} (  t_{kn} ) 
\end{equation}
%%%
where $t_{kn}$ represents the coordinates of the $k$-th surface type on the PC plane ($n$ is either 1 or 2, corresponding to PC 1 and PC 2), i.e., 
%%%
\begin{equation}
s_{kj} = \sum_n t_{kn} V_{nj} + \bar d_j . 
\end{equation}
Assuming that any facet has the same prior probability per area of being a surface type, the prior for each $k$ is:
%%%
\begin{eqnarray}
&& P_{\rm prior} (t_{k1}, t_{k2} ) \, dt_{k1} dt_{k2} \\
&& = \left\{
\begin{array}{ll}
\displaystyle \frac{dt_{k1}  dt_{k2}}{\iint _{\rm allowed} d^2{\bf t}_k} & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_s_range})} \\
0 & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
On the other hand, the likelihood is 
%%%
\begin{eqnarray}
&& P ( U_{in} | t_{kn} ) \\
&& \propto \left\{
\begin{array}{ll}
1 \;\;\; & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_range})} \\
0 \;\;\; & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
Although the graphical meaning of conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) is simply to enclose the data, the judgement is numerically made as follows. 
We first find $\fast $ that satisfies condition (\ref{eq:tilde_f_sum}) from
%%%
\begin{equation}
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
= 
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}
\end{equation}
%%%
or,
\begin{equation}
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
=
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}^{-1}
\label{eq:f=ds-1}
\end{equation}
%%%
Note that the last matrix is a $K\times K$ matrix, and thus inverse is found if $\{ {\bf t}_k \}$ forms a triangle (rather than a straight line or a single dot, in which case $\{ {\bf t}_k \}$ is not a solution anyways).
Then, we ask if the resultant $\fast_{ik}$ are within the range of (\ref{eq:tilde_f_range}).

@@

\memoYF{I still have some concerns about this and the text needs revision. }

@@

We must take caution here that the actual covering fraction of surface types do not necessarily obey gaussian process. 
In this particular case, it appears that the dominant surface type (ocean) is better approximated by gaussian processes.... \memoYF{??}

\memoYF{Have to do with regularization on the maps, but how should we proceed without doing MCMC with both $f_{lk}$ and $s_{kj}$ being fitting parameters? I could operate $W^{+}_{li} \fast_{ik}$ to find $f_{lk}$ with condition (\ref{eq:f_range}) and (\ref{eq:f_sum}); this itself is a bounded linear inversion problem and thus needs some minimization. Also, without regularization, I tend to find scattered $f_{lk}$, perhaps because of the instability of inverse matrix??. }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to EPOXI data}
\label{s:EPOXI}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=\hsize]{raddata_2_PCplane_noreg.pdf}
    \end{center}
    \caption{PC plane from EPOXI observation of June 2008. }
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we apply our analysis procedure to the observed multi-band light curves of the Earth observed by EPOXI mission. 
Specifically we use the same datasets as \citet{Cowan2013}, i.e., the 7-band diurnal light curves observed in June of 2008 \citep{Livengood2011}. 
Using two dominant PCs we extracted, we performed the same analysis described in the previous section
